---
title: "Heuristic Analysis"
author: "Vamshidhar Pandrapagada"
date: "May 16, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Heuristic Analysis

This document compares the different heuristics used in building the agent and highlights reasoning for choosing the heuristic that was ultimately used to submit the agent

## Heuristic 1: Chase Opponent Aggressively
This Heuristic  outputs the score which is equal to the  difference between the number of Player's moves and the 2 times number of opponent's remaining moves.This score chases the opponent aggressively.

Below are the win-rates scored by Heuristic 1 while playing against ID_Improved

```{r echo=FALSE}
library(ggplot2 )
library(gridExtra)
win_rate <- data.frame (
                  Player = c(rep("ID_Improved", 7), rep("Heuristic", 7)),
                  Opponent = c("Random","MM_Open","MM_Center","MM_Improved","AB_Open","AB_Center","AB_Improved"), 
                  Wins  = c(8,6,8,7,5,6,5,10,5,6,7,6,3,4), 
                  Losses = c(2,4,2,3,5,4,5,0,5,4,3,4,7,6)
                )
plot1 <- ggplot(win_rate, aes(x=Opponent, y=Wins, fill=Player)) + geom_bar(stat="identity", position = 'dodge',colour="black") +ggtitle ("                    Win Rate: ID_Improved vs Heuristic 1")
plot2 <- ggplot(win_rate, aes(x=Opponent, y=Losses, fill=Player)) + geom_bar(stat="identity", position = 'dodge',colour="black") +ggtitle ("                    Loss Rate: ID_Improved vs Heuristic 1")
grid.arrange(plot1, plot2, nrow=2)
```
**Win Rate for this heuristic was: 60% **

## Heuristic 2: Evade Corners while nearing End-Game
This heuristic  evaluates the game progress. If the board is occupied at near 70%,  

a. Extract next set of legal moves of the player and the opponent.
b. If the move falls in the corners and board occupancy is  > 70 penalize the move by deducting higher number of points.
c. If the move falls in the corners and board occupancy is  < 70 reward the move by adding lower number of points..
d. Return the sum of differences between the score and the number of moves left for player and the opponent

Below are the win-rates scored by Heuristic 2 while playing against ID_Improved

```{r echo=FALSE}
library(ggplot2 )
library(gridExtra)
win_rate <- data.frame (
                  Player = c(rep("ID_Improved", 7), rep("Heuristic", 7)),
                  Opponent = c("Random","MM_Open","MM_Center","MM_Improved","AB_Open","AB_Center","AB_Improved"), 
                  Wins  = c(8,6,8,7,5,6,5,8,7,8,6,5,5,5), 
                  Losses = c(2,4,2,3,5,4,5,2,3,2,4,5,5,5)
                )
plot1 <- ggplot(win_rate, aes(x=Opponent, y=Wins, fill=Player)) + geom_bar(stat="identity", position = 'dodge',colour="black") +ggtitle ("                    Win Rate: ID_Improved vs Heuristic 2")
plot2 <- ggplot(win_rate, aes(x=Opponent, y=Losses, fill=Player)) + geom_bar(stat="identity", position = 'dodge',colour="black") +ggtitle ("                    Loss Rate: ID_Improved vs Heuristic 2")
grid.arrange(plot1, plot2, nrow=2)
```

**Win Rate for this heuristic was: 63% **

## Heuristic 3: Maximum distance from Walls
This heuristic function evaluates and returns the maximum squared distance between the player and any of the walls.

Below are the win-rates scored by Heuristic 2 while playing against ID_Improved

```{r echo=FALSE}
library(ggplot2 )
library(gridExtra)
win_rate <- data.frame (
                  Player = c(rep("ID_Improved", 7), rep("Heuristic", 7)),
                  Opponent = c("Random","MM_Open","MM_Center","MM_Improved","AB_Open","AB_Center","AB_Improved"), 
                  Wins  = c(8,6,8,7,5,6,5,8,6,7,4,5,5,4), 
                  Losses = c(2,4,2,3,5,4,5,2,4,3,6,5,5,6)
                )
plot1 <- ggplot(win_rate, aes(x=Opponent, y=Wins, fill=Player)) + geom_bar(stat="identity", position = 'dodge',colour="black") +ggtitle ("                    Win Rate: ID_Improved vs Heuristic 3")
plot2 <- ggplot(win_rate, aes(x=Opponent, y=Losses, fill=Player)) + geom_bar(stat="identity", position = 'dodge',colour="black") +ggtitle ("                    Loss Rate: ID_Improved vs Heuristic 3")
grid.arrange(plot1, plot2, nrow=2)
```

**Win Rate for this heuristic was: 58% **

## Observations and Conclusion
**Heuristic number 2: Evade Corners while nearing End-Game** gave consistent results and was able to beat ID_Improved in most of the rounds with the win rate hovering at 72%.
It is very evident that as the game gets near the end state with the board occupied nearly at 70%, avoiding the corners is yielding a higher probability of winning the game.  
I ultimately use Heuristic number 2 for my submission.

